metric-learn: Metric Learning Algorithms in Python

### 2020-07-07 response.tex

\documentclass[a4paper,12pt]{article}

\input{header-PBP.tex}
%\setlength\parindent{0pt}

%%%%%% CROSS-REFERENCING WITH THE PAPER
%\usepackage{zref-xr, zref-user}
%\zexternaldocument*{C:/Users/Emilie/Documents/Main/Recherche/EnCours/S-EmpProcSurSamp/PAPER/ROUND2/EPSS}
% use \zref{} instead of \ref{} to reference the external document(s)

%%%%% DOCUMENT INFORMATIONS

% Authors (short version in brackets, optional)
\author{W. de Vazelhes, CJ Carey, Y. Tang, N. Vauquier, A. Bellet}

% Title of the submitted paper (short version in brackets, optional)
\paper{metric-learn: Metric Learning Algorithms in Python}

% Name of the journal (short version in brackets, optional)
\journal[JMLR]{Journal of Machine Learning Research}

% Submission number
\subnb{19-678(1)}

% Date
\date{\today}

\begin{document}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Description of the revised version}

We thank the referee for his/her helpful comments. As shown by the following
point-by-point response, we have addressed all the comments. Additionally, we have updated the content of the paper to reflect the
improvements made in the last release of \texttt{metric-learn} (v0.6.0, July
2020), in particular the addition of a new algorithm (Sparse Compositional Metric Learning) along with a generic API for triplet-based metric learning algorithms.\footnote{See \url{https://github.com/scikit-learn-contrib/metric-learn/releases} for a full list of changes.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage
\section{Reply to the first referee}

\begin{enumerate}
    \item \textbf{Referee:} There are other metric learning libraries in various programming languages. Perharps a brief comparison with these libraries, specifying similarities and differences, may be of interest to highlight the novelty of this contribution.

    {\it \textbf{Response:} Thank you for the suggestion. We have added a
    paragraph on the positioning with respect to other packages at the end of
    Section~1.}

	\item \textbf{Referee:} Perhaps, as for the proposed code example, I would put one more general example that better captures the essence of what the software is, I see the current example a bit specific and it might be better in the user documentation rather than in the four pages description.

    {\it \textbf{Response:} In the revised version of the paper, we provide
    two separate code snippets. The first one covers supervised metric learning
    and showcases the advantages of our unified interface and tight
    integration with \texttt{scikit-learn} through the use of grid search and
    pipelining with another estimator from \texttt{scikit-learn}. The second
    snippet covers weakly supervised algorithms and is a simplified version of
    the one we had in the initial version of the paper. In both cases, we
    emphasize that the metric learning algorithm used in the snippet can
    be readily switched for another one thanks to our unified interface.}

    \item \textbf{Referee:} It would be interesting too to highlight the advantages of providing a unified interface.

    {\it \textbf{Response:} We have highlighted this better through
    the positioning with respect to other packages and our updated code
    snippets, see above points.}

    \item \textbf{Referee:} Most recent method implemented dates from 2012. Although it is also true that the library implements the best known and possibly most prominent methods, such as LMNN, NCA, MMC or ITML, which are all prior to that year, so I do not see it a drawback. Also in the future work is planned the implementation of more recent methods. In any case, the addition of some more recent method would be positive.

    {\it \textbf{Response:} In the last release of \texttt{metric-learn} 
    (July 2020), we have added a more recent algorithm (Sparse
    Compositional Metric Learning, AAAI 2014) along with a generic API for
    triplet-based metric learning which will make future implementations
    of such algorithms easier. We have updated the content of the paper to
    reflect these changes, see for instance the updated version of Figure~1
    which now illustrates triplet supervision. We will continue to incorporate
    recently proposed algorithms, as emphasized in the conclusion of the paper.}

\end{enumerate}

%\bibliographystyle{plain}
%\bibliography{EPSS}

\end{document}

### 2020-07-27 19-678.tex
Updates from Overleaf

\begin{abstract}%   <- trailing '%' for backward compatibility of .sty file
\texttt{metric-learn} is an open source Python package implementing supervised and weakly-supervised distance metric learning algorithms. As part of \texttt{scikit-learn-contrib}, it provides a unified interface compatible with \texttt{scikit-learn} which allows to easily perform cross-validation, model selection, and pipelining with other machine learning estimators. \texttt{metric-learn} is thoroughly tested and available on PyPi under the MIT licence.
%The source code is available online at \url{http://github.com/scikit-learn-contrib/metric-learn}, as well as the documentation at \url{http://metric-learn.github.io/metric-learn/}.
\end{abstract}

@@

% \william{cf. Nathalie's comment: re-introduce weakly supervised metric laerners}
% Algorithms that learn on pairs or quadruplets take as input 3D array of shape \texttt{(n\_samples, t, n\_features)} (where...). In other words, sampling one element along the first dimension returns one tuple (e.g. a pair), that is a matrix of t lines corresponding to the t elements in the tuple.
% This format is what allows all these API choices, since slicing along the 1st dimension
% While a classical input array \texttt{X} in \texttt{scikit-learn} has shape \texttt{(n\_samples, n\_features)}, an array of tuples has shape \texttt{(n\_tuples, t, n\_features)}, where \texttt{t} is the number of elements in a tuple (e.g. 2 for pairs).
% Note that \texttt{scikit-learn} accepts arrays in a broad sense, in what they call "array-like" objects. \texttt{metric-learn} too accepts "tuple-like" objects.
Weakly supervised algorithms (pair, triplet and quadruplet learners) \texttt{fit} and \texttt{predict} on a set of tuples passed as a 3-dimensional array. Tuples can be pairs, triplets, or quadruplets depending on the algorithm.
Pair learners take as input an array-like \texttt{pairs} of shape \texttt{(n\_pairs, 2, n\_features)}, as well as an array-like \texttt{y\_pairs} of shape \texttt{(n\_pairs,)} giving labels (similar or dissimilar) for each pair.
In order to \texttt{predict} the labels of new pairs, one needs to set a threshold on the distance value.
This threshold can be set manually or automatically calibrated (at fit time or afterwards on a validation set) to optimize a given score such as accuracy or F1-score using the method \texttt{calibrate\_threshold}.
Triplet learners work on array-like of shape \texttt{(n\_triplets, 3, n\_features)}, where for each triplet we want the first element to be closer to the second than to the third one.
Quadruplet learners work on array-like of shape \texttt{(n\_quadruplets, 4, n\_features)}, where for each quadruplet we want the two first elements to be closer together than the two last ones. Both triplet and quadruplet learners can naturally \texttt{predict} whether a new triplet/quadruplet is in the right order by comparing the two pairwise distances.
To illustrate the weakly-supervised learning API, the following code snippet computes cross validation scores for MMC on pairs from Labeled Faces in the Wild \citep{Huang12}. Thanks to our unified interface, MMC can be switched for another pair learner wihtout changing the rest of the code below.
\begin{minted}[%linenos=TRUE,
fontsize=\footnotesize]{python}
from sklearn.datasets import fetch_lfw_pairs 
from sklearn.model_selection import cross_validate, train_test_split 
from metric_learn import MMC
ds = fetch_lfw_pairs()
pairs = ds.pairs.reshape(*ds.pairs.shape[:2], -1)  # we transform 2D images into 1D vectors
y_pairs = 2 * ds.target - 1  # we need the labels to be in {+1, -1}
pairs, _, y_pairs, _ = train_test_split(pairs, y_pairs)
cross_validate(MMC(diagonal=True), pairs, y_pairs, scoring='roc_auc', 
               return_train_score=True, cv=3, n_jobs=-1, verbose=True) 
\end{minted}
%These design choices enable the use of \texttt{scikit-learn}'s scoring functions out of the box, as well as of the standard routines for model selection, including \texttt{GridSearchCV}.

>>>>

\begin{abstract}%   <- trailing '%' for backward compatibility of .sty file
\texttt{metric-learn} is an open source Python package implementing supervised and weakly-supervised distance metric learning algorithms. As part of \texttt{scikit-learn-contrib}, it provides a unified interface compatible with \texttt{scikit-learn} which allows to easily perform cross-validation, model selection, and pipelining with other machine learning estimators. \texttt{metric-learn} is thoroughly tested and available on PyPi under the MIT license.
%The source code is available online at \url{http://github.com/scikit-learn-contrib/metric-learn}, as well as the documentation at \url{http://metric-learn.github.io/metric-learn/}.
\end{abstract}

@@

% \william{cf. Nathalie's comment: re-introduce weakly supervised metric laerners}
% Algorithms that learn on pairs or quadruplets take as input 3D array of shape \texttt{(n\_samples, t, n\_features)} (where...). In other words, sampling one element along the first dimension returns one tuple (e.g. a pair), that is a matrix of t lines corresponding to the t elements in the tuple.
% This format is what allows all these API choices, since slicing along the 1st dimension
% While a classical input array \texttt{X} in \texttt{scikit-learn} has shape \texttt{(n\_samples, n\_features)}, an array of tuples has shape \texttt{(n\_tuples, t, n\_features)}, where \texttt{t} is the number of elements in a tuple (e.g. 2 for pairs).
% Note that \texttt{scikit-learn} accepts arrays in a broad sense, in what they call "array-like" objects. \texttt{metric-learn} too accepts "tuple-like" objects.
Weakly supervised algorithms (pair, triplet and quadruplet learners) \texttt{fit} and \texttt{predict} on a set of tuples passed as a 3-dimensional array. Tuples can be pairs, triplets, or quadruplets depending on the algorithm.
Pair learners take as input an array-like \texttt{pairs} of shape \texttt{(n\_pairs, 2, n\_features)}, as well as an array-like \texttt{y\_pairs} of shape \texttt{(n\_pairs,)} giving labels (similar or dissimilar) for each pair.
In order to \texttt{predict} the labels of new pairs, one needs to set a threshold on the distance value.
This threshold can be set manually or automatically calibrated (at fit time or afterwards on a validation set) to optimize a given score such as accuracy or F1-score using the method \texttt{calibrate\_threshold}.
Triplet learners work on array-like of shape \texttt{(n\_triplets, 3, n\_features)}, where for each triplet we want the first element to be closer to the second than to the third one.
Quadruplet learners work on array-like of shape \texttt{(n\_quadruplets, 4, n\_features)}, where for each quadruplet we want the two first elements to be closer together than the two last ones. Both triplet and quadruplet learners can naturally \texttt{predict} whether a new triplet/quadruplet is in the right order by comparing the two pairwise distances.
To illustrate the weakly-supervised learning API, the following code snippet computes cross validation scores for MMC on pairs from Labeled Faces in the Wild \citep{Huang12}. Thanks to our unified interface, MMC can be switched for another pair learner without changing the rest of the code below.
\begin{minted}[%linenos=TRUE,
fontsize=\footnotesize]{python}
from sklearn.datasets import fetch_lfw_pairs 
from sklearn.model_selection import cross_validate, train_test_split 
from metric_learn import MMC
ds = fetch_lfw_pairs()
pairs = ds.pairs.reshape(*ds.pairs.shape[:2], -1)  # we transform 2D images into 1D vectors
y_pairs = 2 * ds.target - 1  # we need the labels to be in {+1, -1}
pairs, _, y_pairs, _ = train_test_split(pairs, y_pairs)
cross_validate(MMC(diagonal=True), pairs, y_pairs, scoring='roc_auc', 
               return_train_score=True, cv=3, n_jobs=-1, verbose=True) 
\end{minted}
%These design choices enable the use of \texttt{scikit-learn}'s scoring functions out of the box, as well as of the standard routines for model selection, including \texttt{GridSearchCV}.
